### 2020-05-24 norming-study-exp1/run-AMT-experiment/dictionary.tex
started finalizing repository

[{modal:"",id:"1K0",verb:" knows that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2BN",verb:"doesn't believe that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3DP",verb:" doubts that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4R0",verb:" realizes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5AN",verb:"isn't aware that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6CP",verb:" recognizes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7Y0",verb:"is annoyed that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8SN",verb:"isn't sad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9GP",verb:"is glad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10H0",verb:"is happy that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"}],

[{modal:"",id:"1KN",verb:"doesn't know that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2BP",verb:" believes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3D0",verb:" doubts that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4RN",verb:"doesn't realize that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5AP",verb:"is aware that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6C0",verb:" recognizes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7YN",verb:"isn't annoyed that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8SP",verb:"is sad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9G0",verb:"is glad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10HN",verb:"isn't happy that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"}],

[{modal:"Perhaps ",id:"1KP",verb:" knows that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2B0",verb:" believes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3DN",verb:"doesn't doubt that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4RP",verb:" realizes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5A0",verb:"is aware that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6CN",verb:"doesn't recognize that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7YP",verb:"is annoyed that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8S0",verb:"is sad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9GN",verb:"isn't glad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10HP",verb:"is happy that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"}],

[{modal:"",id:"2K0",verb:" knows that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3BN",verb:"doesn't believe that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4DP",verb:" doubts that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5R0",verb:" realizes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6AN",verb:"isn't aware that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7CP",verb:" recognizes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8Y0",verb:"is annoyed that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9SN",verb:"isn't sad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10GP",verb:"is glad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1H0",verb:"is happy that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"}],

[{modal:"",id:"2KN",verb:"doesn't know that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3BP",verb:" believes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4D0",verb:" doubts that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5RN",verb:"doesn't realize that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6AP",verb:"is aware that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7C0",verb:" recognizes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8YN",verb:"isn't annoyed that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9SP",verb:"is sad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10G0",verb:"is glad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1HN",verb:"isn't happy that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"}],

[{modal:"Perhaps ",id:"2KP",verb:" knows that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3B0",verb:" believes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4DN",verb:"doesn't doubt that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5RP",verb:" realizes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6A0",verb:"is aware that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7CN",verb:"doesn't recognize that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8YP",verb:"is annoyed that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9S0",verb:"is sad that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10GN",verb:"isn't glad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1HP",verb:"is happy that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"}],

[{modal:"",id:"3K0",verb:" knows that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4BN",verb:"doesn't believe that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5DP",verb:" doubts that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6R0",verb:" realizes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7AN",verb:"isn't aware that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8CP",verb:" recognizes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9Y0",verb:"is annoyed that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10SN",verb:"isn't sad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1GP",verb:"is glad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2H0",verb:"is happy that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"}],

[{modal:"",id:"3KN",verb:"doesn't know that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4BP",verb:" believes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5D0",verb:" doubts that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6RN",verb:"doesn't realize that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7AP",verb:"is aware that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8C0",verb:" recognizes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9YN",verb:"isn't annoyed that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10SP",verb:"is sad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1G0",verb:"is glad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2HN",verb:"isn't happy that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"}],

[{modal:"Perhaps ",id:"3KP",verb:" knows that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4B0",verb:" believes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5DN",verb:"doesn't doubt that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6RP",verb:" realizes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7A0",verb:"is aware that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8CN",verb:"doesn't recognize that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9YP",verb:"is annoyed that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10S0",verb:"is sad that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1GN",verb:"isn't glad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2HP",verb:"is happy that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"}],

[{modal:"",id:"4K0",verb:" knows that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5BN",verb:"doesn't believe that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6DP",verb:" doubts that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7R0",verb:" realizes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8AN",verb:"isn't aware that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9CP",verb:" recognizes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10Y0",verb:"is annoyed that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1SN",verb:"isn't sad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2GP",verb:"is glad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3H0",verb:"is happy that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"}],

[{modal:"",id:"4KN",verb:"doesn't know that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5BP",verb:" believes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6D0",verb:" doubts that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7RN",verb:"doesn't realize that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8AP",verb:"is aware that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9C0",verb:" recognizes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10YN",verb:"isn't annoyed that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1SP",verb:"is sad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2G0",verb:"is glad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3HN",verb:"isn't happy that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"}],

[{modal:"Perhaps ",id:"4KP",verb:" knows that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5B0",verb:" believes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6DN",verb:"doesn't doubt that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7RP",verb:" realizes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8A0",verb:"is aware that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9CN",verb:"doesn't recognize that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10YP",verb:"is annoyed that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1S0",verb:"is sad that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2GN",verb:"isn't glad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3HP",verb:"is happy that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"}],

[{modal:"",id:"5K0",verb:" knows that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6BN",verb:"doesn't believe that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7DP",verb:" doubts that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8R0",verb:" realizes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9AN",verb:"isn't aware that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10CP",verb:" recognizes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1Y0",verb:"is annoyed that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2SN",verb:"isn't sad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3GP",verb:"is glad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4H0",verb:"is happy that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"}],

[{modal:"",id:"5KN",verb:"doesn't know that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6BP",verb:" believes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7D0",verb:" doubts that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8RN",verb:"doesn't realize that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9AP",verb:"is aware that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10C0",verb:" recognizes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1YN",verb:"isn't annoyed that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2SP",verb:"is sad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3G0",verb:"is glad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4HN",verb:"isn't happy that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"}],

[{modal:"Perhaps ",id:"5KP",verb:" knows that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6B0",verb:" believes that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7DN",verb:"doesn't doubt that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8RP",verb:" realizes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9A0",verb:"is aware that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10CN",verb:"doesn't recognize that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1YP",verb:"is annoyed that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2S0",verb:"is sad that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3GN",verb:"isn't glad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4HP",verb:"is happy that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"}],

[{modal:"",id:"6K0",verb:" knows that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7BN",verb:"doesn't believe that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8DP",verb:" doubts that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9R0",verb:" realizes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10AN",verb:"isn't aware that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1CP",verb:" recognizes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2Y0",verb:"is annoyed that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3SN",verb:"isn't sad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4GP",verb:"is glad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5H0",verb:"is happy that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"}],

[{modal:"",id:"6KN",verb:"doesn't know that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7BP",verb:" believes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8D0",verb:" doubts that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9RN",verb:"doesn't realize that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10AP",verb:"is aware that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1C0",verb:" recognizes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2YN",verb:"isn't annoyed that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3SP",verb:"is sad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4G0",verb:"is glad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5HN",verb:"isn't happy that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"}],

[{modal:"Perhaps ",id:"6KP",verb:" knows that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7B0",verb:" believes that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8DN",verb:"doesn't doubt that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9RP",verb:" realizes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10A0",verb:"is aware that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1CN",verb:"doesn't recognize that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2YP",verb:"is annoyed that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3S0",verb:"is sad that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4GN",verb:"isn't glad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5HP",verb:"is happy that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"}],

[{modal:"",id:"7K0",verb:" knows that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8BN",verb:"doesn't believe that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9DP",verb:" doubts that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10R0",verb:" realizes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1AN",verb:"isn't aware that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2CP",verb:" recognizes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3Y0",verb:"is annoyed that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4SN",verb:"isn't sad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5GP",verb:"is glad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6H0",verb:"is happy that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"}],

[{modal:"",id:"7KN",verb:"doesn't know that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8BP",verb:" believes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9D0",verb:" doubts that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10RN",verb:"doesn't realize that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1AP",verb:"is aware that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2C0",verb:" recognizes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3YN",verb:"isn't annoyed that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4SP",verb:"is sad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5G0",verb:"is glad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6HN",verb:"isn't happy that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"}],

[{modal:"Perhaps ",id:"7KP",verb:" knows that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8B0",verb:" believes that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9DN",verb:"doesn't doubt that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10RP",verb:" realizes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1A0",verb:"is aware that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2CN",verb:"doesn't recognize that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3YP",verb:"is annoyed that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4S0",verb:"is sad that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5GN",verb:"isn't glad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6HP",verb:"is happy that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"}],

[{modal:"",id:"8K0",verb:" knows that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9BN",verb:"doesn't believe that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10DP",verb:" doubts that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1R0",verb:" realizes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2AN",verb:"isn't aware that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3CP",verb:" recognizes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4Y0",verb:"is annoyed mother",name:"Ann",possname:"Ann's",pronoun:"writes",underlined:"poetry"},
{modal:"",id:"5SN",verb:"isn't sad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6GP",verb:"is glad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7H0",verb:"is happy that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"}],

[{modal:"",id:"8KN",verb:"doesn't know that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9BP",verb:" believes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10D0",verb:" doubts that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1RN",verb:"doesn't realize that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2AP",verb:"is aware that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3C0",verb:" recognizes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4YN",verb:"isn't annoyed mother",name:"Ann",possname:"Ann's",pronoun:"writes",underlined:"poetry"},
{modal:"Perhaps ",id:"5SP",verb:"is sad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6G0",verb:"is glad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7HN",verb:"isn't happy that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"}],

[{modal:"Perhaps ",id:"8KP",verb:" knows that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9B0",verb:" believes that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10DN",verb:"doesn't doubt that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1RP",verb:" realizes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2A0",verb:"is aware that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3CN",verb:"doesn't recognize that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4YP",verb:"is annoyed mother",name:"Ann",possname:"Ann's",pronoun:"writes",underlined:"poetry"},
{modal:"",id:"5S0",verb:"is sad that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6GN",verb:"isn't glad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7HP",verb:"is happy that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"}],

[{modal:"",id:"9K0",verb:" knows that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10BN",verb:"doesn't believe that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1DP",verb:" doubts that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2R0",verb:" realizes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3AN",verb:"isn't aware that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4CP",verb:" recognizes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5Y0",verb:"is annoyed that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6SN",verb:"isn't sad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7GP",verb:"is glad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8H0",verb:"is happy that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"}],

[{modal:"",id:"9KN",verb:"doesn't know that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"Perhaps ",id:"10BP",verb:" believes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1D0",verb:" doubts that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2RN",verb:"doesn't realize that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3AP",verb:"is aware that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4C0",verb:" recognizes that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5YN",verb:"isn't annoyed that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6SP",verb:"is sad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7G0",verb:"is glad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8HN",verb:"isn't happy that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"}],

[{modal:"Perhaps ",id:"9KP",verb:" knows that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"},
{modal:"",id:"10B0",verb:" believes that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1DN",verb:"doesn't doubt that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2RP",verb:" realizes that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3A0",verb:"is aware that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4CN",verb:"doesn't recognize that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5YP",verb:"is annoyed that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6S0",verb:"is sad that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7GN",verb:"isn't glad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8HP",verb:"is happy that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"}],

[{modal:"",id:"10K0",verb:" knows that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1BN",verb:"doesn't believe that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"Perhaps ",id:"2DP",verb:" doubts that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3R0",verb:" realizes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4AN",verb:"isn't aware that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"Perhaps ",id:"5CP",verb:" recognizes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6Y0",verb:"is annoyed that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7SN",verb:"isn't sad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"Perhaps ",id:"8GP",verb:"is glad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9H0",verb:"is happy that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"}],

[{modal:"",id:"10KN",verb:"doesn't know that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"Perhaps ",id:"1BP",verb:" believes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2D0",verb:" doubts that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"",id:"3RN",verb:"doesn't realize that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"Perhaps ",id:"4AP",verb:"is aware that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5C0",verb:" recognizes that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"",id:"6YN",verb:"isn't annoyed that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"Perhaps ",id:"7SP",verb:"is sad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8G0",verb:"is glad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"",id:"9HN",verb:"isn't happy that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"}],

[{modal:"Perhaps ",id:"10KP",verb:" knows that",name:"Paul",possname:"Paul's",pronoun:"his",underlined:"karate instructor takes the bus to work"},
{modal:"",id:"1B0",verb:" believes that",name:"Joyce",possname:"Joyce's",pronoun:"her",underlined:"upstairs neighbor owns two dogs"},
{modal:"",id:"2DN",verb:"doesn't doubt that",name:"Jack",possname:"Jack's",pronoun:"his",underlined:"roommate works out six days a week"},
{modal:"Perhaps ",id:"3RP",verb:" realizes that",name:"Kate",possname:"Kate's",pronoun:"her",underlined:"boyfriend lives with his parents"},
{modal:"",id:"4A0",verb:"is aware that",name:"Ann",possname:"Ann's",pronoun:"her",underlined:"mother writes poetry"},
{modal:"",id:"5CN",verb:"doesn't recognize that",name:"Greg",possname:"Greg's",pronoun:"his",underlined:"girlfriend rides horses"},
{modal:"Perhaps ",id:"6YP",verb:"is annoyed that",name:"Eve",possname:"Eve's",pronoun:"her",underlined:"best friend is a waitress"},
{modal:"",id:"7S0",verb:"is sad that",name:"Chad",possname:"Chad's",pronoun:"his",underlined:"soccer coach runs marathons"},
{modal:"",id:"8GN",verb:"isn't glad that",name:"Bea",possname:"Bea's",pronoun:"her",underlined:"ballet teacher is from Russia"},
{modal:"Perhaps ",id:"9HP",verb:"is happy that",name:"Zach",possname:"Zach's",pronoun:"his",underlined:"cousin works for the government"}]